# Reproducible Research Week 2 Project

1. Read activity data in to R, total the number of steps taken each day, and
   convert date column from factor to date variable type
```{r}
activitydata <- read.csv("./activity.csv")
activitydata <- transform(activitydata, date = as.Date(date))
activitytable <- aggregate(activitydata$steps, by=list(activitydata$date), sum)
colnames(activitytable) <- c("date", "TotalStepsTaken")
```

2. Report mean and median of total steps taken each day
```{r}
mean(activitytable$TotalStepsTaken, na.rm=TRUE)
median(activitytable$TotalStepsTaken, na.rm=TRUE)
```

4. Create a histogram of the total number of steps taken each day
```{r}
hist(activitytable$TotalStepsTaken)
```

5. Create a time series plot of the average number of steps taken during each
   interval
```{r}
intervaltable <- aggregate(activitydata$steps, by=list(activitydata$interval), 
                           mean, na.rm=TRUE)
colnames(intervaltable) <- c("interval", "AverageNumberofSteps")
plot(intervaltable$interval, intervaltable$AverageNumberofSteps, type="l")
```

6. Identify interval with the highest average number of steps
```{r}
intervaltable[intervaltable[,2]==max(intervaltable$AverageNumberofSteps), 1]
```

7. Calculate number of rows with missing data
```{r}
table(is.na(activitydata$steps))
```

8. Substitute missing data with average number of steps taken, rounded down, 
   for the recorded interval
```{r}
activitydata[is.na(activitydata$steps),1] <- floor(intervaltable[which(activitydata[is.na(activitydata$steps),3]==intervaltable[1]), 2])
```

9. Create new table and histogram detailing the total number of steps taken
   according to the updated data set
```{r}
activitynoNA <- aggregate(activitydata$steps, by=list(activitydata$date), sum)
colnames(activitynoNA) <- c("date", "TotalStepsTaken")
hist(activitynoNA$TotalStepsTaken)
```

10. Calculate mean and median number of steps taken each day for the new data set
```{r}
mean(activitynoNA$TotalStepsTaken)
median(activitynoNA$TotalStepsTaken)
```

As a result of replacing the missing values with the average number of steps 
taken during the recorded interval, the mean and median number of steps taken
each day both slightly decreased.  

11. Create a new factor variable called DayType that is assigned the value
    "Weekend" if the weekdays() function returns the value "Saturday" or 
    "Sunday" or is assigned the value "Weekday"
```{r}
activitydata$DayType <- factor(NA, c("Weekday","Weekend"))
i=1
for (i in 1:nrow(activitydata)){
      if(weekdays(activitydata[i,2])=="Sunday"||
         weekdays(activitydata[i,2])=="Saturday"){
            activitydata[i,4]<-"Weekend"} 
      
      else {activitydata[i,4] <- "Weekday"}}
```

12. Calculate average number of steps taken for each interval during a "Weekday"
    and a "Weekend"
```{r}
DayTypeTable <- aggregate(activitydata$steps, 
                          by=list(activitydata$DayType,activitydata$interval), 
                          mean)
colnames(DayTypeTable) <- c("DayType", "Interval", "AverageStepsTaken")
```

13. Create a time series plot of the average number of steps taken during each
    measured interval for "Weekday" days and "Weekend" days
```{r}
    library(lattice)
    xyplot(AverageStepsTaken ~ Interval | DayType, data=DayTypeTable, type="l")
```